### 2021-04-21 notes/notes.tex
Added new time table

@@

\section{Motor adaptation}
Motor adaptation to modified environments (e.g. force fields) has been shown to
display the following characteristics:
\begin{enumerate}
\item Savings: Learning a previously-learned environment (after some extinction
phase) is faster than the original learning. This means that the previously
learned model is saved for a while and its learning can be resumed after
extinction. However, \cite{Kojima_Memory_2004} showed that savings can be
``washed out'' if baseline trials (no force field) are inserted between
extinction and the second learning phase.
\item Anterograde interference: Learning an adaptation (e.g. force field
pointing left) slows down learning an opposing adaptation (e.g. force field
pointing right).
\item Rapid de-adaptation: Unlearning an adaptation (even after fully learning
it) is faster than tlearning it \citep{Davidson_Scaling_2004}
\end{enumerate}

>>>>

The previous time table was interrupted by the possibility of the new project. Now that this new project has taken shape, this new time table applies:

\begin{tabularx}{\textwidth}{
  l|
  >{\hsize=0.4\hsize\linewidth=\hsize}X|
  >{\hsize=0.2\hsize\linewidth=\hsize}X|
  >{\hsize=0.2\hsize\linewidth=\hsize}X}
End Date & Result & A priori comments & A posteriori comments \\ \hline \hline
30.04.2021 & Lit. research on savings and deadaptation & Are there many with contextual cues? & \\ \hline
07.05.2021 & Finding model pars to explain observations & This is mostly already done &  \\\hline
14.05.2021 & Looking into error-clamp behavior & Maybe there's something to it. If there is, an extra week should be allocated to simulations. &  \\\hline
--- & Start writing the paper. & & \\ \hline

\end{tabularx}
\end{chapter}

@@

\section{Notes from }
There is extensive evidence that fast trial-by-trial error-based learning relies on the cerebellum. Patients with cerebellar lesions show substantial impairment in fast adaptation across many task domains51–55

\section{Motor adaptation}
In typical motor adaptation experiments \citep[e.g.][]{Medina_Mechanism_2001,
  Davidson_Scaling_2004,Kojima_Memory_2004}, the goal of the participant is to
move some squishy appendage (e.g. eyes, hand) from point A to point B,
repeatedly, under different circumstances. These experiments consist of a
number of phases in different order (and some of them not present in some
experiments). These phases are the following:
\begin{enumerate}
\item Baseline (O): Participants move their appendage from A to B without being
hindered, as they normally would in their life.
\item Adaptation (A, B, ...): In this phase, the circumstances in which the movement
happens (still from A to B) are changed. For example, in human experiments with
arm movement, the arm will be subjected to forces by an external actor (e.g. a
mechanical arm attached to the participant's hand). The participant's goal is
to learn to counteract these forces and perform the movement from A to B as
well as in Baseline.
\item De-adaptation: In this phase, the circumstances are changed, either
reverted to Baseline or to some other circumstance. For example, the forces
applied by the mechanical arm are removed, or their direction is changed to the
opposite.
\item Error-clamp (E): In this phase, the error in the movement is forced to
zero. For example, the mechanical arm applies a force that forces the
participant's hand to move in a straight line from A to B, even if the
participant applies perpendicular forces.
\end{enumerate}

Using these phases, several phenomena has been shown:
\begin{enumerate}
\item Savings: Learning a previously-learned environment (after some extinction
phase) is faster than the original learning. This means that the previously
learned model is saved for a while and its learning can be resumed after
extinction. However, \cite{Kojima_Memory_2004} showed that savings can be
``washed out'' if baseline trials (no force field) are inserted between
extinction and the second learning phase; their findings are perhaps confounded
by the passage of time, as this inserted baseline makes a lot of time pass.
\item Anterograde interference: Learning an adaptation (e.g. force field
pointing left) slows down learning an opposing adaptation (e.g. force field
pointing right).
\item Rapid de-adaptation: Unlearning an adaptation (even after fully learning
it) is faster than learning it \citep{Davidson_Scaling_2004}
\item Spontaneous recovery: After a de-adaptation phase, an error-clamp phase
is introduced and participants show a rapid adaptation in the direction of A,
which slowly goes back to zero across hundreds (?) of trials.
\end{enumerate}

The experiment carried out with monkeys by \cite{Kojima_Memory_2004} used
different structures (see their figure 1C). For example, to show savings they
used O-A-B-A, where in A and B the saccade target moved during the saccade to
make the monkey think that their saccades were off. During A and B, opposite
directions were used (e.g. in A the monkeys had to learn to increase their
gain, and in B learn to decrease it). They found savings in the form of a quick
learning period during the second A block that took the monkey close to the
adaptation at the end of the first A block (see their figure 3A); after that,
learning proceeded as before.

Using an O-A-B-O-A paradigm, they showed that savings were no longer
present. On the other hand, using an O-A-B-E-A paradigm (where the error-clamp
phase consisted on putting the monkey in the dark) they found savings (their
figures 5A and 7A).

\cite{Smith_Interacting_2006} showed spontaneous recovery in a O-A-B-E
paradigm. From their data it is unclear what exactly happens during B, but
during E it seems like participants adapted a bit in the direction of B and
then quickly jumped to the direction of A, from which they slowly went to zero.

Similarly, \cite{Ethier_Spontaneous_2008} who spontaneous recovery with
saccades in an E-A-B-E paradigm. From their results (e.g. their figure 2) it
seems possible that some participants go back to A's adaptation and some others
to O's. Additionally, they introduced 30s breaks during adaptation and showed
that these breaks had a ``forgetting'' effect, i.e. adaptation was set back by
a small amount after each break. The further posit that introducing such a
break before the second E phase would eliminate spontaneous recovery entirely.

\cite{Huang_Persistence_2009} showed that if participants are placed in an
adaptation environment where the force gradually ramps up (across many trials),
participants show slower de-adaptation during an error-clamp phase, compared to
participants in which the force was suddenly increased during adaptation. They
also showed that participants alter their learning rates to better fit the
volatility of the environment, regardless of the direction of adaptation.
\end{chapter}
